Prey movement shapes the development of predator expertise in a virtual bi-trophic system

### 2024-07-10 manuscript/manuscript.Rmd
manuscript completed

In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*. First, we investigate the development of expertise at the population level. We hypothesize that the predator population's success will increase with experience up to a certain level where it will stabilize [@dukas2019]. However, we expect this pattern to change depending on the movement of the prey encountered. We hypothesize that prey will influence the development of expertise, and predict that faster prey will reduce the gain in expertise. Therefore, we then investigate how prey movement influences the development of expertise at the individual level. If prey speed does not influence hunting success, we predict that the gain in expertise will be similar among individuals. Alternatively, if prey speed influences hunting success, then the acquisition of expertise will vary among individuals.

@@

*revoir la première phrase, devrait intégrer les proies*

Predators usually increase prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023]. Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise because of the challenges of investigating direct predator-prey interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience. Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the movement of the prey encountered.

Our results suggest that the acquisition of hunting expertise was honed through extensive practice. The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated. Our observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019]. However, the prey's speed was important in mediating this pattern at the population level, probably because prey also increased their speed as predators gained experience. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022]. This would explain why the relationship was concave in the model where we did not account for prey speed, because adept predators were encountering prey with higher speeds. Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.

Prey speed also mediated differences among predator players in the development of their expertise, implying that prey antipredator behaviour can impair the development of an individual predator's expertise. Individual predator players probably differed in their capacity to adjust to difficult prey. For example, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011]. Thus, predators that couldn't develop counter-strategies for faster prey were likely at a disadvantage.

*ici, répéter les études utilisant le camouflage, ensuite inclure la phrase des compromis, et conclure avec une avenue évolutive.*

**Such trade-offs may reflect limitations in learning all the skills required to successfully hunt all types of prey [@healy1992;@belisle1997;@dukas2019], particularly if the skills required to hunt slower prey are nontransferable to faster prey.**

However, even after we took the prey's speed into account, there were still important differences in expertise acquisition among predator players, suggesting that other antipredator tactics were potentially involved. For example, the prey can use camouflage and hide to avoid detection (not currently measurable in *DBD*), requiring predators to learn how to exploit visual cues that facilitate cryptic prey detection, such as habitat characteristics or prey colour patterns [@ehlinger1989;@hughes.etal1992;@warburton2003;@szopa-comley.etal2020]. Furthermore, it is hypothesized that longer time intervals between hunting events could hinder or delay the acquisition of expertise as individuals may forget information with longer delays [@endler1991;@wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions.

**Of course, the life of neither the predator nor the prey players are at stake, such that emerging patterns could be driven by their motivation to win and not "true" survival. Yet,**
**A potential caveat is that the more flexible hunters might have experimented with various tactics out of boredom, which could impede ecologically realistic interpretations of our data. However, the consistent association between this tactic and heterogeneous prey groups gives us confidence that it emerged from their interactions. One limitation of our study was that we couldn't monitor all the matches of the prey, which prevented us from assessing their responses to the predator through their experience.** Hence, future studies should aim at monitoring reciprocal behavioural dynamics through time, which may reveal important insights on the mechanisms driving stable equilibria in predator-prey systems.

**Comment by reviewer:**

Some characteristics to be aware of: 
- Predators do not compete.
- Both predators and prey can fail in a round yet still carry that experience and knowledge on to the next round with no real penalty (think of the movie Groundhog Day).
- Predator and prey are surely playing a game with each other, but prey may also be playing a game among themselves, which may affect their game with the predator. 
- This particular video game appears to have a few unusual characteristics.  Prey do not appear to forage for resources. but rather they create a public good.  Furthermore, handling time for the predator is extremely complicated and lengthy.
- Finally, it may be rather complicated to discern all of the rules of how the different virtual environments work and what  all the aptitudes of all the different predators and prey are.

@@

We found support of our prediction that prey were driving individual differences in expertise in a predator population. Our results suggest that predators learned with experience, as their success increased...  Even though individuals were not all equally successful. Lastly, virtual systems are increasingly recognized as being useful to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.

>>>>

In this study, we assess how repeated encounters with prey shapes predator hunting success using data from players in *Dead by Daylight*. First, we investigate how the predator population develops its hunting expertise. We hypothesize that the predator population's success will increase with experience up to a certain level where it will stabilize [@dukas2019]. However, we expect this pattern to change depending on the movement of the prey encountered. We hypothesize that prey will influence the development of expertise, and predict that faster prey will reduce the gain in expertise. Therefore, we then investigate how prey movement influences the development of expertise at the individual level. If prey speed does not influence hunting success, we predict that the gain in expertise will be similar among individuals. Alternatively, if prey speed influences hunting success, then the acquisition of expertise will vary among individuals.

@@

Predators usually optimize prey consumption through time by acquiring hunting expertise [@dukas2019;@wooster.etal2023]. Yet, few studies have empirically tested how prey antipredator strategies mediate the acquisition of hunting expertise due to the challenges of investigating direct predator-prey interactions in the wild. By capitalizing on a virtual predator-prey system where interactions were directly monitored, we found that a predator-player population in *Dead by Daylight* increased their hunting success with experience. Yet, there were important differences among individuals in the development of said expertise, which we found to be in part due to the movement of the prey encountered.

Our results suggest that the acquisition of hunting expertise was honed through extensive practice. The predator population displayed an asymptotic relationship between experience and success, wherein initial gains in success were significant but gradually stabilized as experience accumulated. Our observations are consistent with empirical studies of expertise in both humans and nonhuman animals [reviewed in @dukas2019]. However, the prey's speed was important in mediating this pattern at the population level. Encountering faster prey led to more difficult encounters for predators. We previously showed in *DBD* that faster movement is an effective strategy used by the prey to avoid predation [@fraserfranco.etal2022], and other studies have found that as well [@walker.etal2005;@kelley.magurran2011;@martin.etal2022]. This could explain why the relationship was concave in the model where we did not account for prey speed, because the prey were probably increasing their speed as they gained experience. Because the matchmaking algorithm pairs players based on their skill, adept predators were probably encountering prey moving at higher speeds. Thus, our results suggest that predators can gain expertise and maintain success when they encounter prey that move at speeds lower or close to the population-average.

Prey speed also mediated differences among predator players in the development of their expertise. Individual predator players probably differed in their capacity to adjust to difficult prey. Animals are expected to have limited attention that constrain diet choice and image search formation [@dukas.etal2001]. Moreover, hunting faster prey requires specialized cognitive abilities and coordination that are energetically costly [@kelley.magurran2011]. Thus, predators that couldn't develop counter-strategies to detect or chase faster prey were likely at a disadvantage. Parallel observations have been outlined in studies of prey camouflage strategies. For example, @troscianko.etal2018 showed in a computer experiment involving humans that disruptive camouflage was efficient at preventing the acquisition of expertise during search image formation. Human subjects exposed to a restricted set of strategies were also less efficient compared to those exposed to a variety of strategies. Therefore, our observations suggest that prey antipredator behaviour can also impair predator expertise acquisition, with potential implications for predator-prey dynamics.

Despite adjusting for the prey's speed, discernible variations in expertise acquisition among predator players persisted, indicating the possible influence of other factors. It is hypothesized that longer time intervals between hunting events can hinder or delay the acquisition of expertise because individuals may forget information when delays are longer [@endler1991;@wright.etal2022]. In *DBD*, a predator that played 300 matches in the span of six months might forget more critical information related to prey detection or escape patterns than one that played 300 matches in the span of six days. Investigating the impact of such time lags in future analyses may reveal important insights on the outcome of predator-prey interactions. Another limitation of our study is that the life of neither the predator nor the prey players are at stake in *DBD*, such that emerging patterns could be driven by their motivation to win and not "true" survival. For example, some players could experiment with the game out of boredom, which could impede ecologically realistic interpretations of our data. Lastly, the prey can be consumed in a match but carry that experience on to the next round without any penalty. Thus, interpretations of our results must be made with these caveats in mind.

@@

We found support of our hypothesis that prey antipredator behaviour was driving individual differences in expertise in a predator population. Our results suggest that predators gained expertise, as their success increased wit extensive practice, but that prey speed impaired expertise acquisition. Future analyses should investigate how efficient antipredator tactics are associated with prey experience, as this may reveal important insights on the eco-evolutionary dynamics of predator prey interactions. Lastly, virtual systems are increasingly used as systems to test hypotheses on ecological interactions [@beauchamp2020;@cere.etal2021;@fraserfranco.etal2022;@lymbery.etal2023;@santostefano.etal2024]. We therefore hope that our study will inspire more collaborations between scientists and the videogame industry to tackle fundamental questions in ecology.
